### 2025-12-10 submissions/03-jpam-RR/manuscript_main.tex
submit R&R!

Asian Americans represent the fastest-growing racial group in the United States, yet their experiences with discrimination and identity formation remain underexplored in economic research.\footnote{The 2020 Census counted more than 20 million Asian Americans---6.4 percent of the population—nearly double the number counted two decades earlier \autocite{floodsarahIntegratedPublicUse2021a}. The Asian American population numbers are based on the author's calculations from the Current Population Survey and US Census data.} Unlike other minority groups, Asian Americans occupy a distinctive position in America's racial hierarchy—simultaneously experiencing discrimination and being labeled as "perpetual foreigners" while being characterized through the "model minority" stereotype. This dual status creates complex incentives around racial identity choices that fundamentally differ from other groups' experiences, as Asian racial identity can signal both academic excellence and perpetual foreignness.

@@

Various contextual factors, including anti-Asian sentiment and stereotype threat, can influence how individuals navigate their racial identity choices. Recent events have brought renewed attention to how external hostility shapes Asian American experiences \autocite{gover2020anti}. In this paper, I examine the determinants of Asian racial identity reporting and analyze how Asian Americans strategically select between Asian and White racial identities. Specifically, I investigate how anti-Asian bias, education, and family income shape decisions to identify racially as Asian American. I also break down the analysis by generation and family structure (interracial versus endogamous parents) and examine what racial identities individuals with objective Asian ancestry report (Asian only, White only, multiracial, etc.).

@@

Measuring identity choices outside of lab settings is challenging, requiring both objective ancestry indicators and subjective identity measures. I leverage birthplace and ancestry data to construct objective Asian ancestry measures, then analyze deviations between objective ancestry and subjective racial identity. The analysis reveals that racial identity reporting responds to both individual characteristics and environmental factors reflecting local discrimination levels.

@@

The analysis by family structure reveals additional heterogeneity: bias effects are strongest among children from mixed-race families, with a one standard deviation bias increase correlating with a 15 percentage point decrease in Asian racial identity among children of Asian fathers and White mothers, and a 10 percentage point decrease among children of White fathers and Asian mothers. Among adult samples, these patterns are even more pronounced, with second-generation adults from White father-Asian mother families showing 24 percentage point decreases in Asian racial identity in response to bias increases.

@@

Multinomial logit analyses reveal that anti-Asian bias fundamentally reshapes racial identity choices. I find that when bias increases from its lowest to highest levels, individuals dramatically shift away from ``Asian only'' racial identity, with probabilities decreasing from 98\% to 48\% across all generations, while probabilities of choosing ``White only'' racial identity rise from 1\% to 43\%. In contrast, I find that parental education produces modest and inconsistent effects that vary by family composition: maternal college education increases Asian identity reporting by approximately 11 percentage points among second-generation children of Asian father-White mother families but shows minimal effects in White father-Asian mother families, while paternal education effects remain small across both family types. These patterns suggest that anti-Asian bias, rather than socioeconomic factors, primarily drives identity choices, causing research using subjective measures to misestimate Asian-White gaps in highly prejudiced areas.

@@

Recognizing the identity flexibility that characterizes Asian American experiences, I investigate the economic determinants driving racial self-identification decisions. \textcite{hadah2024hispanicidentity} finds that bias and self-reported Hispanic identity are negatively associated among objectively Hispanic immigrants. I examine how personal and environmental factors influence the complexity of endogenous racial identity among Asian Americans, recognizing that the model minority stereotype creates unique economic incentive structures not present for other groups. The empirical analysis documents how observable characteristics---individual traits and societal attitudes---affect racial identity reporting among Asian Americans. These findings have important implications for measuring racial economic disparities and understanding how discrimination operates in modern labor markets.

@@

My data shows consistent Asian identity reporting regardless of whether mother (72\%), father (72\%), or child/other caregiver (87\%) serves as respondent, as shown in Table \ref{tab:hispbyproxy}.\footnote{According to CPS guidelines, household respondents must be at least 15 years old with sufficient household knowledge. When the proxy is `self,' the respondent ranges from 15 to 17 years old.} Since my analysis compares high and low bias states, estimates remain valid provided reporting patterns don't systematically differ between these contexts. Furthermore, ethnic attrition patterns among adults align with those observed in children, as shown in Table (\ref{tab:hispbygen}) for children versus Table (\ref{tab:hispbygen-adults}) for adults, suggesting that proxy reporting aligns with individuals' self-identification.

@@

I construct anti-Asian sentiment measures using implicit association tests, American National Election Studies, and hate crimes targeting Asian Americans from 2004--2021. The implicit association test measures conceptual associations---for example, linking Asian Americans with negative stereotypes---and evaluative responses. Respondents rapidly categorize words into screen-displayed categories. Figure (\ref{fig:iatexamples}) shows examples from Harvard's Project Implicit skin tone test.

@@

I develop another racial animus proxy using ANES surveys \autocite{anes2021} measuring discrimination against racial groups. ANES, conducted since 1948, enjoys widespread political science usage. The survey examines attitudes toward different racial groups, voting intentions, and political questions. I employ several 2004--2020 ANES questions measuring racial animus. The racial animus index averages responses across multiple animus-measuring questions.\footnote{Questions parallel those used by \textcite{charlesPrejudiceWagesEmpirical2008}: (1) ``Conditions Make it Difficult for Blacks to Succeed'', (2) ``Blacks Should Not Have Special Favors to Succeed'', (3) ``Blacks Must Try Harder to Succeed'', (4) ``Blacks Gotten Less than They Deserve Over the Past Few Years'', and (5) ``Feeling Thermometer Toward Asians.''} While the ANES racial animus questions primarily focus on attitudes toward Black Americans, research demonstrates that racial prejudices are highly correlated across different minority groups, with individuals who express bias toward one racial minority typically holding similar attitudes toward others \autocite{almasalkhi2023links, mora2020antiblackness}. These measures therefore capture broader patterns of racial animus that extend beyond anti-Black sentiment specifically. When combined with the Asian-focused IAT measures and hate crimes against Asian Americans in my composite index, this multi-proxy approach weights the bias measure more heavily toward Asian-specific prejudice while still capturing the general racial climate.

@@

Third-generation patterns reveal increasing identity fluidity (Figure \ref{fig:histogram-thirdgen}). White only identity becomes most common overall (35\%), followed by Asian only (29\%). The number of Asian grandparents strongly predicts choices: those with one Asian grandparent report 54\% White only, while those with four Asian grandparents report 92\% Asian only.

@@

Results by family structure (Figure \ref{plot01-regression-byparent}) reveal stronger bias effects among children from interracial families. Among second-generation children, a one standard deviation increase in anti-Asian bias correlates with statistically insignificant 5 percentage point decreases for those with endogamous Asian parents (panel B), but significant 15 percentage point decreases for Asian father-White mother children (panel C) and 10 percentage point decreases for White father-Asian mother children (panel D). Maternal college education consistently shows positive effects on Asian identity reporting in endogamous families and Asian father-White mother families, with particularly strong associations (approximately 11 percentage points) in the latter group.

@@

In this section, I discuss the results from the multinomial logit analysis. These results reveal how anti-Asian bias differentially affects the probability of choosing ``Asian only,'' ``White only,'' or ``Asian and White'' racial identity, with particularly pronounced effects among mixed-race Asian Americans. Given that the results are mainly driven by individuals with mixed ancestry, I focus the discussion on overall patterns and subsamples with mixed ancestry. The multinomial logit coefficients are in log-odds form, which are difficult to interpret directly. Therefore, I present predicted probabilities to facilitate interpretation.

\textbf{Overall Patterns Among All Generations.} Figure \ref{fig:pp-all-gen} shows predicted probabilities from the multinomial logit model across all generations. Anti-Asian bias produces the most pronounced effects on racial identity choice. When bias increases from its lowest level ($-0.7$) to its highest level ($1.3$), the predicted probability of reporting ``Asian only'' identity decreases dramatically from 98\% to 48\%, while ``White only'' identity increases from 1\% to 43\%, and ``Asian and White'' identity shows a modest increase from 1\% to 8\%. In contrast, gender and parental education produce minimal effects on racial identity choices, with male and female respondents showing virtually identical probabilities across all identity categories, and parental college education generating small, inconsistent effects.

\textbf{Second-Generation Subsample by Parental Composition.} Parental composition reveals substantial differences between Asian father-White mother (AW) and White father-Asian mother (WA) families (Figures \ref{fig:pp-second-aw} and \ref{fig:pp-second-wa}). Among AW families, as bias increases from minimum to maximum levels ($-0.7$ to $1.3$), ``Asian only'' identity decreases from 66\% to 28\%, while ``White only'' increases from 3\% to 55\%. Among WA families, ``Asian only'' decreases from 54\% to 18\%, ``Asian and White'' increases from 40\% to 62\%, and ``White only'' increases from 6\% to 20\%. Gender shows minimal effects in both family types. Parental education effects differ by family structure: among AW families, maternal college education increases ``Asian only'' identity from 51\% to 62\%, while among WA families, parental education produces virtually no change across identity categories.

Adult second-generation results (Figure \ref{fig:pp-secondgen-adults-aw-wa}) show similar patterns. Among AW families (Panel \ref{subfig:pp-secgen-aw-bias}), as bias increases from minimum to maximum levels, ``White only'' identity increases from 12\% to 39\%, while ``Asian and White'' decreases from 46\% to 16\%. Among WA families (Panel \ref{subfig:pp-secgen-wa-bias}), ``White only'' increases from 27\% to 38\%, ``Asian and White'' increases from 25\% to 40\%, and ``Asian only'' decreases from 47\% to 22\%. Gender shows minimal effects on identity choices for both family types.

\textbf{Third-Generation Subsample by Grandparent Composition.} Third-generation identity choices vary systematically with the number of Asian grandparents (Figures \ref{fig:pp-third-one}, \ref{fig:pp-third-two}, and \ref{fig:pp-third-three}). Anti-Asian bias produces dramatically different effects across ancestral compositions. Those with one Asian grandparent show modest responses to bias, with ``Asian only'' decreasing from 11\% to 3\% and ``Asian and White'' increasing from 34\% to 48\% as bias moves from minimum to maximum. Those with two Asian grandparents show stronger effects, with ``Asian only'' decreasing from 48\% to 13\% and ``White only'' increasing from 21\% to 49\%. Those with three Asian grandparents maintain high ``Asian only'' identification ($\approxeq 95\%$) until bias reaches its highest levels, when it drops 20 percentage points.

Gender effects are minimal across all grandparent types, with males and females showing similar identity distributions. Parental education operates differently by ancestral composition: maternal college education has modest effects among those with one grandparent, strongest effects among those with two grandparents (increasing ``Asian only'' from 18\% to 23\%), and significant effects among those with three grandparents (increasing ``Asian only'' from 88\% to 96\%).

@@

I also investigate relationships between anti-Asian bias and migration. Since CPS doesn't report birth states, I use 2004--2021 Censuses to construct second-generation Asian American samples \autocite{floodsarahIntegratedPublicUse2021}. I construct mover variables indicating whether second-generation Asian Americans moved from birth states to other states. I use the following models to estimate relationships between anti-Asian bias and migration:

@@

I show the results of estimating Equations (\ref{eq:migration-3}), (\ref{eq:migration-4}), and (\ref{eq:migration-5}) in Table (\ref{regtab-mig-01}) columns (1), (2), and (3), respectively. Among second-generation immigrants, no significant correlations exist between anti-Asian bias and migration decisions. Among second-generation Asian American movers, those self-reporting Asian racial identity live in states with 0.06 standard deviations greater bias than their birth states. While this result shows selection into more biased states among Asian-identifying second-generation immigrants, it doesn't affect the main results showing correlations between anti-Asian bias and Asian racial identity reporting. Since Asian-identifying individuals are moving to higher-bias states, my assessment of the relationship between bias and Asian racial identity reporting might underestimate bias effects.

Several concerns merit discussion. First, CPS self-reported identity comes from household respondents---parents or adult caregivers. I view parent- or caregiver-reported identity as an accurate representation of children's identity since parents essentially shape children's self-concepts. Moreover, since my analysis compares high- and low-bias states, estimates remain valid provided reporting patterns don't systematically differ between these contexts.

Moreover, \textcite{duncanIntermarriageIntergenerationalTransmission2011} show that reported Hispanic racial identity doesn't vary with household respondent identity. Consistent with this, Table (\ref{tab:hispbyproxy}) shows that Asian racial identity reporting equals 72 percentage points when mothers or fathers serve as proxies and 87 percentage points when children or other caregivers serve as household respondents.\footnote{According to CPS guidelines, household respondents must be at least 15 years old with sufficient household knowledge. When the proxy is `self,' respondents range from 15 to 17 years old.} To further address this concern, I examine adult Asian American samples where individuals self-report racial identity. I find similar patterns of ethnic attrition and bias effects among adults (Table \ref{tab:hispbygen-adults} and Figures \ref{plot01-regression-gen-adults}--\ref{plot01-regression-byparent-adults}).

Another concern involves reverse causality between larger Asian American or Black populations in states and bias levels. Greater Asian American populations might affect resident bias levels. To demonstrate this isn't occurring, Figure (\ref{scatter-plot-1}) plots self-reported Asian American state percentages against average anti-Asian bias in those states. I find no correlations between anti-Asian bias and Asian American state populations, making reverse causality unlikely.

Finally, bias and Asian racial identity reporting relationship estimates could be biased if non-Asian-identifying individuals migrate to more prejudiced states. I've shown above this isn't occurring (Table \ref{regtab-mig-01}). I find no evidence of relationships between migration decisions and anti-Asian bias. Additionally, those reporting Asian racial identity moved from less biased birthplaces and lived in more biased states at survey times. Thus, my results might underestimate relationships between anti-Asian bias and Asian racial identity reporting.

@@

Using multinomial logit analysis, I show that anti-Asian bias fundamentally reshapes racial identity. When bias increases from minimum to maximum levels, the probability of reporting ``Asian only'' identity decreases dramatically from 98\% to 48\%, while ``White only'' identity increases from 1\% to 43\%. These strategic identity choices are most pronounced among mixed-race families, where high bias environments drive substantial shifts toward White racial identity among Asian father-White mother families and toward multiracial identity among White father-Asian mother families.

@@

This study encourages further research into relationships between bias and self-reported identities for other groups. Similar analysis could examine bias effects on sexual minority identities and other ethnic and racial minorities such as Black, Native American, and Arab American populations in the United States. Researchers could also explore outcome differences between ethnic and racial minorities who self-report versus those who don't using restricted administrative data.

The research opens several avenues for future investigation. First, scholars could examine how recent anti-Asian violence following COVID-19 has influenced identity patterns, providing natural experiments in bias effects. Second, researchers might explore identity choices in specific institutional contexts like college admissions or workplace advancement, where model minority stereotypes create complex incentive structures. Third, analysis could extend to other Asian American subgroups, recognizing that Chinese, Korean, Vietnamese, and other communities face distinct stereotypes and discrimination patterns. Fourth, more granular geographic analysis at the county or zip-code level could provide more precise measures of local bias, though data limitations prevent this in the current analysis.

>>>>

Asian Americans represent the fastest-growing racial group in the United States, yet their experiences with discrimination and identity formation remain underexplored in economic research.\footnote{The 2020 Census counted more than 20 million Asian Americans---6.4 percent of the population—nearly double the number counted two decades earlier \autocite{floodsarahIntegratedPublicUse2021a}. The Asian American population numbers are based on the author's calculations from the Current Population Survey and US Census data.} Unlike other minority groups, Asian Americans occupy a distinctive position in America's racial hierarchy—simultaneously experiencing discrimination and being labeled as ``perpetual foreigners'' while also being characterized through the `model minority' stereotype \autocite{foukaImmigrantsAmericansRace2022}. This dual status creates complex incentives around racial identity choices that fundamentally differ from other groups' experiences. Understanding how both anti-Asian bias and socioeconomic status influence racial identity decisions is crucial, as the selection could systematically distort the true enumeration of Asian in the US, the empirical assessments of Asian American outcomes and the true extent of discrimination they face.

@@

Various contextual factors, including anti-Asian sentiment and stereotypes, can influence how individuals navigate their racial identity choices. The COVID-19 pandemic have brought renewed attention to how external hostility shapes Asian American experiences \autocite{gover2020anti}. In this paper, I examine the determinants of Asian racial identity reporting and analyze how Asian Americans strategically select between Asian and White racial identities. Specifically, I investigate how anti-Asian bias, education, and family income shape decisions to identify racially as Asian American. I also break down the analysis by generation and family structure (interracial versus endogamous parents) and examine what racial identities individuals with objective Asian ancestry report (Asian only, White only, multiracial, etc.).

@@

Measuring identity choices outside of lab settings is challenging, requiring both objective ancestry indicators and subjective identity measures. I leverage birthplace and ancestry data to construct objective Asian ancestry measures, then analyze deviations between objective ancestry and subjective racial identity. I find that racial identity reporting responds to both individual characteristics and environmental factors reflecting local discrimination levels.

@@

When analyzing the heterogeneity results by family structure, I find that: bias effects are strongest among children from mixed-race families. A one standard deviation bias increase leads to a 15 percentage point decrease in Asian racial identity among children of Asian fathers and White mothers and a 10 percentage point decrease among children of White fathers and Asian mothers. Among adult samples, these patterns are even more pronounced, with second-generation adults from White father-Asian mother families showing 24 percentage point decreases in Asian racial identity in response to bias increases.

@@

Multinomial logit analyses reveal that anti-Asian bias fundamentally reshapes racial identity choices. Among adult Asian father-White mother respondents, a one standard deviation increase in bias shifts reporting toward ``White only'' (about +0.14) and away from multiracial identity (about --0.11); White father-Asian mother adults move out of ``Asian only'' (about --0.13) into multiracial (about +0.09) and slightly ``White only'' (about +0.05). Third-generation respondents with two Asian grandparents similarly shift away from ``Asian only'' (about --0.10) toward both alternatives (``White only'' about +0.05; ``Asian and White'' about +0.06), while bias effects are small and imprecise when only one grandparent is Asian. Parental education exerts modest, composition-specific effects—e.g., maternal education boosts ``Asian only'' for some groups and paternal education can reduce ``White only''—but these shifts are much smaller than the bias-driven movements. These patterns underscore that local prejudice, more than socioeconomic resources, drives strategic identity choices and can distort measured Asian-White gaps in highly prejudiced areas.

@@

Recognizing the identity flexibility that characterizes Asian American experiences, I investigate the economic determinants driving racial self-identification decisions. I examine how personal and environmental factors influence the complexity of endogenous racial identity among Asian Americans, recognizing that the model minority stereotype creates unique economic incentive structures not present for other groups. The empirical analysis documents how observable characteristics---individual traits and societal attitudes---affect racial identity reporting among Asian Americans. These findings have important implications for measuring racial economic disparities and understanding how discrimination operates in modern labor markets.

@@

My data shows consistent Asian identity reporting regardless of whether mother (72\%), father (72\%), or child/other caregiver (87\%) serves as respondent, as shown in Table \ref{tab:hispbyproxy}.\footnote{According to CPS guidelines, household respondents must be at least 15 years old with sufficient household knowledge. When the proxy is `self,' the respondent ranges from 15 to 17 years old. Since my analysis compares high and low bias states, estimates remain valid provided reporting patterns do not systematically differ between these contexts.} Furthermore, ethnic attrition patterns among adults align with those observed in children, as shown in Table (\ref{tab:hispbygen}) for children versus Table (\ref{tab:hispbygen-adults}) for adults, suggesting that proxy reporting aligns with individuals' self-identification.

@@

I construct anti-Asian sentiment measures using implicit association tests, American National Election Studies, and hate crimes targeting Asian Americans from 2004--2021.

The implicit association test measures conceptual associations---for example, linking Asian Americans with negative stereotypes---and evaluative responses. Respondents rapidly categorize words into screen-displayed categories. Figure (\ref{fig:iatexamples}) shows examples from Harvard's Project Implicit skin tone test.

@@

Using the ANES survey, I develop another racial animus proxy using ANES surveys measuring discrimination against racial groups \autocite{anes2021}. ANES, conducted since 1948, is widely used in political science research. The survey examines attitudes toward different racial groups, voting intentions, and political questions. I employ several 2004--2020 ANES questions measuring racial animus. The racial animus index averages responses across multiple animus-measuring questions.\footnote{Questions parallel those used by \textcite{charlesPrejudiceWagesEmpirical2008}: (1) ``Conditions Make it Difficult for Blacks to Succeed'', (2) ``Blacks Should Not Have Special Favors to Succeed'', (3) ``Blacks Must Try Harder to Succeed'', (4) ``Blacks Gotten Less than They Deserve Over the Past Few Years'', and (5) ``Feeling Thermometer Toward Asians.''} While the ANES racial animus questions primarily focus on attitudes toward Black Americans, research demonstrates that racial prejudices are highly correlated across different minority groups, with individuals who express bias toward one racial minority typically holding similar attitudes toward others \autocite{almasalkhi2023links, mora2020antiblackness}. These measures therefore capture broader patterns of racial animus that extend beyond anti-Black sentiment specifically. When combined with the Asian-focused IAT measures and hate crimes against Asian Americans in my composite index, this multi-proxy approach weights the bias measure more heavily toward Asian-specific prejudice while still capturing the general racial climate.

@@

Third-generation patterns reveal increasing identity fluidity (Figure \ref{fig:histogram-thirdgen}). White only identity becomes most common overall (35\%), followed by Asian only (29\%), Asian and White/Pacific Islander (29\%), and other multiracial identities (7\%). The number of Asian grandparents strongly predicts choices: those with one Asian grandparent report 54\% White only, while those with four Asian grandparents report 92\% Asian only.

@@

Results by family structure (Figure \ref{plot01-regression-byparent}) reveal stronger bias effects among children from interracial families. Among second-generation children, a one standard deviation increase in anti-Asian bias correlates with statistically insignificant 5 percentage point decrease for those with endogamous Asian parents (panel B), but significant 15 percentage point decrease for Asian father-White mother children (panel C) and 10 percentage point decrease for White father-Asian mother children (panel D). Maternal college education consistently shows positive effects on Asian identity reporting in endogamous families and Asian father-White mother families, with particularly strong associations (approximately 11 percentage points) in the latter group.

@@

This section presents results from the multinomial logit analysis, revealing how anti-Asian bias differentially affects the probability of choosing ``Asian only,'' ``White only,'' or ``Asian and White'' racial identity. The effects prove particularly pronounced among mixed-race Asian Americans, on whom I focus the discussion. Because multinomial logit models express coefficients in log-odds form, I present average marginal effects showing how each explanatory variable affects the probability of choosing each racial identity, as displayed in Figures \ref{fig:marginal-effects-second-parental}--\ref{fig:marginal-effects-third-grandparental}.\footnote{I present the predicted probabilities from the multinomial logit regressions in the online appendix in Figures \ref{fig:pp-second-aw}--\ref{fig:pp-third-two}.}

\textbf{Second-Generation Subsample by Parental Composition.} Parental composition reveals substantial differences between Asian father-White mother (AW) and White father-Asian mother (WA) families (Figure \ref{fig:marginal-effects-second-parental}). Among AW families, an increase of one standard deviation in anti-Asian bias from the mean increases the probability of reporting ``Asian only'' identity by 13 percentage points and decreases the probability of reporting ``White only'' identity by 20 percentage points. Among WA families, a one standard deviation increase in bias decreases the probability of reporting ``Asian only'' by 12 percentage points and increases the probability of reporting ``White only'' identity by 10 percentage points.

Gender exerts minimal effects in both family types. Parental education operates differently by family structure. Among AW families, maternal college education increases the probability of reporting ``Asian only'' and decreases the probability of reporting ``White only.'' Paternal college education increases the probability of reporting ``Asian and White'' and decreases the probability of reporting ``Asian only.'' Among WA families, maternal education produces virtually no change across identity categories, while paternal college education increases the probability of reporting ``Asian and White'' identity and decreases the probability of reporting ``White only'' identity.

Adult second-generation marginal effects (Figure \ref{fig:marginal-effects-second-adults}) show similar directional patterns but with wide confidence intervals. Among AW adults, higher anti-Asian bias pushes identity toward ``White only'' (about +0.14) and away from ``Asian and White'' (about --0.11), with a small, imprecise decline for ``Asian only.'' Years of education increase the probability of reporting ``Asian only'' (roughly +0.02) and reduce ``White only'' reporting (about --0.02), while household income shows near-zero effects across categories. Gender effects remain modest: females tilt slightly toward ``Asian and White'' with offsetting small declines in the other categories.

Among WA adults, anti-Asian bias markedly reduces ``Asian only'' identification (about --0.13) and shifts probability toward multiracial (about +0.09) and, to a lesser extent, ``White only'' (about +0.05), though estimates are imprecise. Education again nudges responses toward ``Asian only'' (about +0.02) and away from ``White only'' (about --0.02), with negligible change in multiracial reporting. Household income reduces ``Asian only'' (about --0.02) and modestly raises ``Asian and White'' identity (about +0.02). Gender effects are small and statistically indistinguishable from zero.

\textbf{Third-Generation Subsample by Grandparent Composition.} Third-generation identity choices vary systematically with the number of Asian grandparents (Figure \ref{fig:marginal-effects-third-grandparental}). Anti-Asian bias has small, imprecise effects for those with one Asian grandparent, tilting probability modestly toward multiracial identity (about +0.04) and away from both ``Asian only'' and ``White only'' (about --0.02 each). With two Asian grandparents, higher bias more clearly pulls respondents away from ``Asian only'' (about --0.10) toward both ``White only'' (+0.05) and ``Asian and White'' (+0.06).

Parental education matters differently by grandparent composition. Among those with one Asian grandparent, maternal college completion shifts identity toward ``Asian and White'' (about +0.09) and away from ``White only'' (about --0.10), with little effect on ``Asian only.'' Paternal college shows the same pattern but slightly smaller magnitudes (``Asian and White'' about +0.07; ``White only'' about --0.03; ``Asian only'' about --0.03). Among those with two Asian grandparents, maternal college markedly raises ``Asian only'' (about +0.07) and lowers ``Asian and White'' (about --0.07), with negligible change in ``White only.'' Paternal college reduces ``White only'' (about --0.11) and increases multiracial reporting (about +0.12), leaving ``Asian only'' unchanged. Gender effects remain small in both groups.

Taken together, the multinomial results show that anti-Asian bias moves identity away from ``Asian only'' toward ``White only'' or multiracial options, especially among mixed-ancestry adults and third-generation respondents with two Asian grandparents. Education and income partially offset this drift for some groups (e.g., AW adults, WA adults, and third-generation individuals with two Asian grandparents), while parental education often shifts mixed-ancestry respondents toward maintaining at least partial Asian identification. This selective retention underscores that studies relying on self-reported racial identity may overstate Asian American socioeconomic success and understate assimilation speed, as individuals with more resources and education are more likely to keep Asian identification visible in the data.

@@

I also investigate relationships between anti-Asian bias and migration. Since CPS does not report birth states, I use 2004--2021 Censuses to construct second-generation Asian American samples \autocite{floodsarahIntegratedPublicUse2021}. I construct mover variables indicating whether second-generation Asian Americans moved from birth states to other states. I use the following models to estimate relationships between anti-Asian bias and migration:

@@

I show the results of estimating Equations (\ref{eq:migration-3}), (\ref{eq:migration-4}), and (\ref{eq:migration-5}) in Table (\ref{regtab-mig-01}) columns (1), (2), and (3), respectively. Among second-generation immigrants, no significant correlations exist between anti-Asian bias and migration decisions. Among second-generation Asian American movers, those self-reporting Asian racial identity live in states with 0.06 standard deviations greater bias than their birth states. While this result shows selection into more biased states among Asian-identifying second-generation immigrants, it does not affect the main results showing correlations between anti-Asian bias and Asian racial identity reporting. Since Asian-identifying individuals are moving to higher-bias states, my assessment of the relationship between bias and Asian racial identity reporting might underestimate bias effects.

Several concerns merit discussion. First, CPS self-reported identity comes from household respondents---parents or adult caregivers. I view parent- or caregiver-reported identity as an accurate representation of children's identity since parents essentially shape children's self-concepts. Moreover, since my analysis compares high- and low-bias states, estimates remain valid provided reporting patterns do not systematically differ between these contexts.

Moreover, \textcite{duncanIntermarriageIntergenerationalTransmission2011} show that reported Hispanic racial identity does not vary with household respondent identity. Consistent with this, Table (\ref{tab:hispbyproxy}) shows that Asian racial identity reporting equals 72 percentage points when mothers or fathers serve as proxies and 87 percentage points when children or other caregivers serve as household respondents.\footnote{According to CPS guidelines, household respondents must be at least 15 years old with sufficient household knowledge. When the proxy is `self,' respondents range from 15 to 17 years old.} To further address this concern, I examine adult Asian American samples where individuals self-report racial identity. I find similar patterns of ethnic attrition and bias effects among adults (Table \ref{tab:hispbygen-adults} and Figures \ref{plot01-regression-gen-adults}--\ref{plot01-regression-byparent-adults}).

Another concern involves reverse causality between larger Asian American or Black populations in states and bias levels. Greater Asian American populations might affect resident bias levels. To demonstrate this is not occurring, Figure (\ref{scatter-plot-1}) plots self-reported Asian American state percentages against average anti-Asian bias in those states. I find no correlations between anti-Asian bias and Asian American state populations, making reverse causality unlikely.

Finally, bias and Asian racial identity reporting relationship estimates could be biased if non-Asian-identifying individuals migrate to more prejudiced states. I have shown above this is not occurring (Table \ref{regtab-mig-01}). I find no evidence of relationships between migration decisions and anti-Asian bias. Additionally, those reporting Asian racial identity moved from less biased birthplaces and lived in more biased states at survey times. Thus, my results might underestimate relationships between anti-Asian bias and Asian racial identity reporting.

@@

Using multinomial logit analysis, I show that anti-Asian bias fundamentally reshapes racial identity. A one standard deviation increase in bias nudges adult Asian father-White mother respondents toward ``White only'' identity (about +0.14) and away from multiracial identity (about --0.11), while White father-Asian mother adults exit ``Asian only'' (about --0.13) toward multiracial (about +0.09) and, to a lesser extent, ``White only'' (about +0.05). Among third-generation respondents with two Asian grandparents, higher bias shifts reporting away from ``Asian only'' (about --0.10) toward both ``White only'' (+0.05) and ``Asian and White'' (+0.06), with much smaller, imprecise changes when only one grandparent is Asian. These strategic identity adjustments highlight that bias pressures are most pronounced for mixed-ancestry families and later-generation individuals with greater identity flexibility.

@@

This study encourages further research into relationships between bias and self-reported identities for other groups. Similar analysis could examine bias effects on sexual minority identities and other ethnic and racial minorities such as Black, Native American, and Arab American populations in the United States. Researchers could also explore outcome differences between ethnic and racial minorities who self-report versus those who do not using restricted administrative data.

The research opens several avenues for future investigation. First, scholars could examine how recent anti-Asian violence following the COVID-19 pandemic has influenced identity patterns, providing natural experiments in bias effects. Second, researchers might explore identity choices in specific institutional contexts like college admissions or workplace advancement, where model minority stereotypes create complex incentive structures. Third, analysis could extend to other Asian American subgroups, recognizing that Chinese, Korean, Vietnamese, and other communities face distinct stereotypes and discrimination patterns. Fourth, more granular geographic analysis at the county or zip-code level could provide more precise measures of local bias, though data limitations prevent this in the current analysis.
